Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach

### 2018-09-03 Manuscript.Rmd
Many changes to figures, methods, and ensuring that the new ROIs work

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  pdf_document: default
  word_document: default
  html_document: default
bibliography: mPFC_paper.bib
---

@@

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = 5,564 per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of the correlations so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and $W$ is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (**Add to results**), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

@@

The following shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning.

```{r Template partition, out.width="60%", fig.align = "center", fig.cap = "Figure 2. Brain partition for an example subject (100307). Fiedler vector values (top) are mapped onto the brain, dividing it into positive and negative communities. The bottom brain shows the binarized Fiedler vector, with yellow areas denoting the DMN community (as indicated by coverage of area 7m)."}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)

>>>>

---
title: 'Parsing mPFC: A Joint Meta-Analytic and Graph Theoretic Approach'
author: "Claudio Toro-Serey & Joseph T. McGuire"
output:
  html_document: default
  word_document: default
  pdf_document: default
bibliography: mPFC_paper.bib
---

@@

``` {r For HCP data, fig.align="center", echo=FALSE, message = FALSE}
setwd('./For_HCP')
temp <- list.files(pattern = '*Binary*')
FB <- lapply(temp, read.table)
Summed <- do.call(cbind, lapply(FB, "[[", 'V1')) - 1
FB_Means <- rowMeans(Summed)

@@

To establish each subject's network, we selected all the vertices contained within the ROIs (n = **5,564 (UPDATE)** per subject) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, where surface vertices were the nodes, and edges the correlations among all of them. Next, we applied Fisher's r to z transformation, and took the exponential of the correlations so that all weights were positive while maintaining the shape of the original correlation distribution, and retained the edges whose adjusted p-values remained significant after correction for multiple tests (FDR < 0.05). As part of the evaluation step, this procedure (and the community detection outlined below) was applied at four levels: 1) each session separately; 2) the concatenated timeseries from each pair of daily sessions; 3) each subject's dataset; and 4) on each step of a sliding window analysis (see Evaluation for more details).

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlations with every other vertex), and $W$ is the correlation matrix. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler Vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected (Fiedler, 1975). Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community as the one containing the one DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks (Fortunato & Hric, 2016).

@@

```{r Density calculation, out.width="60%", fig.align = "center"}
# Load density data
setwd('./Densities/')
temp <- list.files()
Densities <- do.call(rbind, lapply(temp, read.csv))

# Get the mean density for all data, daily scans, and sessions
# Check distributions, since median might be better
mdAll <- mean(Densities$Density[grep("Overall", Densities$Type)])
sddAll <- sd(Densities$Density[grep("Overall", Densities$Type)])
mdDay <- mean(Densities$Density[grep("Day", Densities$Type)])
mdSess <- mean(Densities$Density[grep("Session", Densities$Type)])
```

First, we estimated each individual's DMN coverage using the whole time series (4800 TRs). On average, networks had high densities (i.e. highly connected) regardless of the amount of data to produce them (mean = `r round(mdAll, digits = 2)`, SD = `r sddAll`). Figure 2 shows a stereotypical subdivision of the ROIs for a single subject (100307) using spectral partitioning.

```{r Template partition, out.width="60%", fig.align = "center", fig.cap = "Figure 2. Brain partition for an example subject (100307). Fiedler vector values (top) are mapped onto the brain surface, dividing it into positive and negative communities. The bottom brain shows the binarized Fiedler vector, with yellow areas denoting the DMN community (as indicated by coverage of area 7m)."}
# Load data
setwd('./Summary/')
temp <- list.files()
subjList <- sapply(temp, substring, first=1, last=6)
Summaries <- lapply(temp, read.csv)
nSubj <- length(Summaries)
